---
title: "Residual Analysis"
author: "Sandeep Anand"
date: "April 7, 2017"
output: html_document
---

```{r setup and global options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Analysis of the diamond data Set in R

```{r plot,highlight=TRUE, echo=FALSE, background="#F6F7F6", fig.height=5, fig.width=5}

library(ggplot2)
library(UsingR)

data(diamond)

g = ggplot(diamond, aes(x=carat, y=price))
g <- g + xlab("Carats (weight)")
g <- g + ylab("Price (Value in DLR)")
g <- g+ geom_point(size=6, color = "black", alpha=0.2)
g <- g+ geom_point(size=5, color = "blue", alpha=0.2)
g <- g+geom_smooth(method = "lm", color = "brown")
g
```

---
## Fitting the Linear regression model

### **This will be printing the coefficients beta0, beta1 for the linear regression model**

```{r fit, echo=TRUE}
fit<-lm(price ~ carat, data = diamond)
coef(fit)
```


---
## Mean Centering the Predictor variable to make more sense of the coefficients 

### **This will be printing the coefficients beta0, beta1 for the linear regression model with the mean centered**

+ Getting the value in carat/10

```{r fit2, echo=TRUE}
fit2<-lm(price ~ I((carat-mean(carat))*10), data = diamond)
coef(fit2)
```

---
## Predicting with some new Data 

+ Using the Predict funtion 

```{r fit3, echo=TRUE}
newx <- c(0.34, 0.45, 0.78)
predict(fit, newdata = data.frame(carat = newx))
```


---
## Illustrating the Residual variation throgh a residual plot

+ Using the Predict funtion 
+ Residuals have to sum to 0 and so they have to lie above and below the line y=0

```{r Residuals, echo=TRUE}
data(diamond)
y<-diamond$price; x<-diamond$carat; n<-length(y)

#Creating the linear regression object that is created by lm
fitval<-lm(y~x)


#Summary of Fit and elements of fit
summary(fitval)

e<-resid(fitval)

#Vector of the Prediction of the observed x values ie, carat values 
yhat<-predict(fitval)

# Finding the maximum of the residual variation through the formula as well as manually
max(abs(e-(y-yhat)))

max(abs(e-(y-coef(fitval)[1] - coef(fitval)[2]*x)))

#Plot-1
plot(y, x, xlab = "Carats (Weight)", ylab = "Price ($)", bg="lightblue", col="black", cex=1.1, pch=21, frame=FALSE, title(main = "Residuals seen in terms of the linear fit regression line"))
abline(fitval,lwd=2)
for(i in 1:n)
  lines(c(x[i], x[i]), c(y[i], yhat[i]), col="purple", lwd=2)

#Plot-2
plot(x, e, xlab = "Carats (Weight)", ylab = "Price ($)", bg="lightblue", col="black", cex=2, pch=21, frame=FALSE, title(main = "Residual True representation"))
abline(h=0, lwd=2)
for(i in 1 :n)
  lines(c(x[i],x[i]), c(e[i], 0), col="red", lwd=2)

# Sum of Residuals 
sum(e)

# Sum of Residuals multiplied with the predictors
sum(e*x)
```


---
## Illustrating another example through Linear fit and Residual analysis

+ It is not necessary that Linear Fit model will always be the best fit for a model, but it will still be important in representing important information
+ Residuals have to sum to 0 and so they have to lie above and below the line y=0

```{r Residual of a linear/sine,echo=TRUE, fig.height=5, fig.width=8}
x<-runif(100, -3,3)
y<-x+sin(x)+rnorm(100,sd=0.2)

library(ggplot2)

g<-ggplot(data.frame(x=x, y=resid(lm(y~x))),aes(x=x,y=y))
g<-g+ geom_hline(yintercept = 0, size=2)
g<-g+ geom_point(size=7, color="black", alpha=0.4)
g<-g+ geom_point(size=5, color="red", alpha=0.4)
g<-g + xlab("X") + ylab("Residual")

g

```

---
## Heteroskedsticity

+ The variability of the data incresing along the horizontal line , y=0 , as x increases 
+ This porperty can be seen clearly with the help of a residual graph

```{r Diamond Data residual plot,echo=TRUE,fig.height=4.5, fig.width=4.5}
e<-c(resid(lm(price ~ 1, data=diamond)), resid(lm(price~carat, data=diamond)))

fitval1<-factor(c(rep("Itc", nrow(diamond)),
                  rep("Itc, Slope", nrow(diamond))))

library(ggplot2)

g<-ggplot(data.frame(e=e, fit=fitval1),aes(y=e, x=fitval1, fill=fitval1))
g<-g+ geom_dotplot(binaxis = "y", size=2, stackdir = "center", binwidth = 20)
g<-g + xlab("Fitting approach") + ylab("Residual Price")

g

```


