---
title: "GLM - Generalized Linear Models"
author: "Sandeep Anand"
date: "April 22, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Generalized Linear Models

  - We will use logistic regression as below when we would consider a Binomial outcome
  - On the Logit scale you would like to see if the coefficient is close to 0 or not , on the exponential scale that should be 1

```{r Ravens Logistic Regression}
#path = getwd()
#path_for_file = paste(path, "/RavensData/",sep="")
#dir.create(path_for_file)
#-- download.file() not working , pleae chek
#download.file("https://dl.dropbox.com/u/7710864/data/ravensData.rda",destfile="./RavensData/ravensData.rda",method="curl")
load("./RavensData/ravensData.rda")
head(ravensData)
logRegRavens<-glm(ravensData$ravenWin ~ ravensData$ravenScore, family = "binomial")
summary(logRegRavens)
```

## Fitten Curves  
  * Only a part of the S curve is shown below
  * Component of the S curve where the data is present was restricted 
  * Took the X values , multiplied it with the coefficients , added the intercept to it and then got the ratio of exp(val)/(1+exp(val)) and that is the probability ie, on the Y       asis
  * equation block:
  $$P_r= \frac{\mathrm{e}^{p}}{(1+\mathrm{e}^{p})}$$

```{r FItted Curve, echo=FALSE}
plot(ravensData$ravenScore, logRegRavens$fitted, pch=19, col="blue", xlab = "Score", ylab = "Prob Ravens Win")
```

## Understanding the coefficients
  * 11% increase in the probability of winning of every point that the ravens score

### Sometool like Anova is very useful in porting a factor variable in and out of a model , when there a number of factor variables available

```{r exponents, echo=TRUE}
exp(logRegRavens$coeff)
exp(confint(logRegRavens))
anova(logRegRavens, test="Chisq")
```

## Interpreting Odds Ratios

  - Not probabilities
  - Odds ratio of 1 = no difference in odds
  - Log odds ratio of 0 = no difference in odds
  - Odds ratio < 0.5 or > 2 commonly a "moderate effect"
  - Relative risk $\frac{\rm{Pr}(RW_i | RS_i = 10)}{\rm{Pr}(RW_i | RS_i = 0)}$ often easier to interpret, harder to estimate
  - For small probabilities RR $\approx$ OR but they are not the same!
  
## Modelling Count Data - Poisson Distribution

  - Poisson Distribution is useful model for counts and rates
  - Modelling web traffic rates
  - Contigency tables data
  - Poisson Distribution becomes very much like a normal distributin as $$t\lambda$$, increases
  - Mean and Variance of a Poisson Distribution are equal
  - We are going to assume that our data is poisson distributed with a mean and that mean takes this form - difference between Linear Model
  - Everything is log() , we log the mean of the Outcome, and this is considered as one of the best transformations we can do 
  - Slope coeff. is the relative increase or decrease in the mean per one unit chage in the regressor.
  - For these type of modelling it is important that variance also increases with variance , and that can be checked using a Residual vs. Fitted curve 
  - Also This can be achieved by the use of quasi-poisson models, to hold the right trend
  - Check the use ofsandwich variance estimator, made famous by generalized estimating equations
  - One problem that happens a lot with poisson distribution is when 0's occur a lot , and that is called 0 inflation, there are packages for handling zero inflation
  
```{r Poisson Distribution, echo=TRUE}
par(mfrow=c(1,3))
plot(0:10, dpois(0:10, lambda = 2), type = "h", frame.plot = TRUE)
plot(0:20, dpois(0:20, lambda = 10), type = "h", frame.plot = TRUE)
plot(0:300, dpois(0:300, lambda = 100), type = "h", frame.plot = TRUE)
```
