---
title: "Plotly demo using IMDB kaggle dataset"
author: "Sandeep Anand"
date: "June 11, 2017"
output: 
  ioslides_presentation:
    widescreen: true
    smaller: true
    fig_width: 5
    fig_height: 3
    fig_caption: true
params:
  n: 100
  d: !r Sys.Date()
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

## R Markdown

###link used for the data 
  + https://www.kaggle.com/deepmatrix/imdb-5000-movie-dataset/downloads/movie_metadata.csv.zip

## Getting the Data & Data Cleaning

### Points of Interest 
  - Based on the IMDB score(>=7), check the budget spent on the movie , considering a USA movie on recent years [2014-2016]
  - Based on the IMDB Score(>6), check the number of movies released on different separate genres considering a USA movie on recent years [2013-2016]

```{r cars, echo = FALSE, warning=FALSE}
library(tools)
path_orginal <- "C:\\Public\\R-Projects\\R-projects\\datasets"
path_store <- paste(path_orginal, "movie_metadata.csv", sep = "\\")

for (i in list.files(path = path_orginal)) {
  if (file_ext(i) == "csv" && i=="movie_metadata.csv"){
    df = read.csv(file = path_store, header = TRUE, sep=",")
  }
}

#head(df)
predictors<-names(df)


```

## Plot the number of movies released in the US between 2013 to 2016 based on separate Genres and imdb score > 6

```{r imdb score vs Language Category, echo=FALSE, warning=FALSE, fig.align='center', fig.height=4,fig.width=6}
library(plotly)
set.seed(1234)
touse_b1 <- c("genres", "movie_title","gross","country","title_year","imdb_score")
newdf_b1 <- df[touse_b1]

newdf_b1_filt1 <- newdf_b1[which((newdf_b1$title_year >= 2014) & (newdf_b1$country == "USA") & (newdf_b1$imdb_score>6)), ]

# Making sense of the movie title
newdf_b1_filt1$movie_title<-as.character(newdf_b1_filt1$movie_title)
newdf_b1_filt1$movie_title <- substr(newdf_b1_filt1$movie_title, 1, nchar(newdf_b1_filt1$movie_title)-2)

newdf_b1_filt1$genres<-as.character(newdf_b1_filt1$genres)

# Get all the Genres categories 
gencat_filt1<-strsplit(newdf_b1_filt1$genres, "\\|")
gencat_un<-unique(gencat_filt1)

mylistofcat <- list()

for (i in 1:length(gencat_un)) {
    mylistofcat[i] <- gencat_un[[i]]
}

mylistofuniquecat <- unique(mylistofcat)

lhs <- c()

#Create a data frame here
for (i in 1:length(mylistofuniquecat)) 
  {
    lhs[i] <- mylistofuniquecat[[i]]
    assign(lhs[i], 0)
  }

df_final <- data.frame(lhs)
df_final$count <- rep(0, length(mylistofuniquecat))
df_final$lhs <- as.character(df_final$lhs)

#Now to increment the count of the particular genre as you see in the main "imdb", data frame
library(Hmisc)
for (i in 1:length(gencat_filt1)) {
  val<-match(gencat_filt1[[i]], df_final$lhs) #This gives the indices where the match occurs
  for (j in 1:length(val)){
    inc(df_final$count[val[j]])<-1
  }
}
library(plotly)
plot_ly(data = df_final, x= ~lhs, y= ~count, type = 'bar', text=text, 
        marker = list(color = 'rgb(158, 202, 225)',
                      line = list(color = 'rgb(8,50,118)',
                                  width = 1.5)))

```


```{r imdb score Vs the Budget, echo=FALSE, warning=FALSE}
library(plotly)

touse <- c("movie_title","budget","title_year","imdb_score","country")
newdf<- df[touse]

#subset based on Year > 2014 only

newdf_final <- newdf[ which((newdf$title_year >= 2014) & (newdf$country == "USA") & (newdf$imdb_score >= 7) & (!(is.na(newdf$imdb_score)))), ]
newdf_final$budget <- newdf_final$budget/10^6
newdf_final$movie_title<-as.character(newdf_final$movie_title)
#strsplit(newdf_final$movie_title, "")
#newdf_final$movie_title <- substr(newdf_final$movie_title,1,length(newdf_final))
splitpoint <- lapply(strsplit(newdf_final$movie_title[1], ""), tail, 2)
#as.character(splitpoint[[1]])

newdf_final$movie_title <- substr(newdf_final$movie_title,1,nchar(newdf_final$movie_title)-2)
```

## Plotting Budget(in million) Based on imdb Score

```{r plot 1, echo=FALSE, warning=FALSE}
library(ggplot2)
library(plotly)
library(ggforce)

#Draw each page
  
p <-ggplot(data = newdf_final, aes(x=budget, y=imdb_score)) +
    geom_point(aes(text = paste("Movie Title", movie_title))) +
    geom_smooth(aes(colour = title_year, fill=title_year)) + facet_wrap_paginate(~title_year, ncol = 2, nrow = 2, page = 1) +
    labs(x = "Budget(in mill US)", y = "imdb Score")

ggplotly(p)

```