---
title: "Instacart Buyer Analysis"
author: "Sandeep Anand"
output: html_notebook
---

#Description 

In this competition, Instacart is challenging the Kaggle community to use this anonymized data on customer orders over time to predict which previously purchased products will be in a user's next order. 

## Getting the Data
```{r Getting the Data, eval=TRUE, warning=FALSE, echo=FALSE}
path1 <- "C:/Public/Instacart/"
for (i in list.files(path=path1, pattern = "\\.zip$")){
  unzip(i, overwrite = TRUE)
  }

```

```{r Data2, eval=TRUE, warning=FALSE, echo=FALSE}
names <- character(0)
for (j in list.files(path = path1, pattern = "\\.csv$")) 
  {
    names<-c(names,unlist(strsplit(j,"\\."))[1])
}

```

```{r Data3, eval=TRUE, warning=FALSE, echo=FALSE}
filenames<-list.files(path = path1, pattern = "*.csv$", full.names = TRUE)
ldf <- lapply(filenames, read.csv)
```

```{r assignment, eval=TRUE, warning=FALSE, echo=FALSE}
for(i in 1:(length(names)-1)){
  nam <- paste(names[i], "df", sep = ".")
  assign(nam, ldf[[i]])
}
```

## Assignments to be used for Data-frame names:
  - "aisles.df"
  - "departments.df"
  - "order_products__prior.df"
  - "order_products__train.df"
  - "orders.df" 

```{r Orders.df Proc, eval=TRUE, warning=FALSE, echo=FALSE}
order_train.df<-orders.df[orders.df$eval_set=="train",]
order_prior.df<-orders.df[orders.df$eval_set=="prior",]
order_test.df<-orders.df[orders.df$eval_set=="test",]

head(order_products__prior.df)

head(order_prior.df)
```

## Getting the mapping done For Prior set
  - Get the mapping between order_id and the associated products for it for a particular user
  - New Dataframes for getting 
    + user-id, Order-id and count of products for each of the order-ids
    + splitting the data-frame that corresponds to orders.csv (filtered based on prior) into multiple sub datafrmaes, based on user_id
    + Get the products for each Order for each of the above data frames and put them into a list that will be an extra row
    + Helper functions are present to break the number of userIds(~206k) into 207 chunks of 1000 each to process them faster
    + chks is a List of length 1 which has another list of length 207
    + Never create multiple variables: not be a wise decision to create multiple vars , why not process and get the output in this function itself

```{r Helper Functions, eval=TRUE, warning=FALSE, echo=FALSE}
require(plyr)
chunks <- function(d, n){      
    chunks <- split(d, ceiling(seq_along(d)/n))
    names(chunks) <- NULL
    return(chunks)
}

chunksplyr <- function(d, n) {
  is<-seq(from=1, to=length(d), by=n)
  if (tail(is,1)!=length(d)) {
    is<-c(is, length(d))
  }
  
  chunks<-ddply(head(seq_along(is),-1),
                function (i) {
                  start <- is[i];
                  end <- is[i+1]-1;
                  d[start:end]
                }
                )
  lc <- length(chunks)
  td<-tail(d,1)
  chunks[[lc]] <- c(chunks[[lc]], td)
  return(chunks)
}

```


```{r mapping, eval=TRUE, warning=FALSE, echo=FALSE}
temp1 <- c() 
listofdfs <- c()
numberobs = length(order_prior.df$order_id)
numberuserids = unique(order_prior.df$user_id)
order_prior.df_cols<-names(order_prior.df)
order_prior.df_sub = order_prior.df[order_prior.df_cols[1:2]]

cat("The total number of userIds:", length(numberuserids))
#Break the data frame order_prior.df into multiple dfs based on user_id

system.time(chks <- chunks(numberuserids,100))
length(chks)
#system.time(chks2 <- chunksplyr(numberuserids,1000))

df = data.frame("user_id"=integer(), "order_id"=integer(), "product_count"=integer())

x <- chks[[1]][1:100]

idval <- integer()
order <- integer()
pcount <- integer()

system.time(
  lapply(x, function(x){
    for (i in seq_along(x)) {
      tempval <- order_prior.df[order_prior.df$user_id == x[i],]$order_id
      lapply(seq_along(tempval), function(j){
        idval[j] <<- x[i]
        order[j] <<- tempval[j]
        pcount[j] <<- length(which(order_products__prior.df$order_id == tempval[j]))
      })
    }
    temp <- data.frame("user_id"=idval, "order_id"=order, "product_count"=pcount)
    df <<- rbind(df, temp)
  })
)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).
