---
title: "Linear Regression"
author: "Sandeep Anand"
date: "March 31, 2017"
output: html_document
---

```{r setup, include=FALSE, highlight=TRUE, tidy=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## Demonstation of the lm() function and Simple Linear Regression

### This part is contributed from data archive at Princeton Examples from http://data.princeton.edu/R/readingData.html 

```{r Reading the Data}
fpe<-read.table("http://data.princeton.edu/wws509/datasets/effort.dat", header = TRUE)
```

## Scatter Plot of Change based on Effort of the above Family Planning data


```{r chagne vs Effort, echo=FALSE}
plot(fpe$effort, fpe$change, pch=21, bg="green")
title("Scatterplot of change by Effort", col.main="#3366CC")
```

## Fitting a linear Model

> Note first that lm is a function, and we assign the result to an object that we choose to call lmfit (for linear model fit). This stores the results of the fit for later examination.

> The argument to lm is a model formula, which has the response on the left of the tilde ~ (read "is modeled as") and a Wilkinson-Rogers model specification formula on the right. R uses

  +	to combine elementary terms, as in A+B
  + :	for interactions, as in A:B
  *	for both main effects and interactions, so A*B = A+B+A:B


```{r fitting a linear model, echo=TRUE}
lmfit<-lm( change ~ setting+effort, data = fpe)
summary(lmfit)
anova(lmfit)
plot(lmfit)
```

```{r Using Galton data to find the relations, fig.height=4,fig.width=4, echo=TRUE}
library(UsingR) 
library(ggplot2)
data(galton)

y<-galton$child #Outcome
x<-galton$parent #Predictor

coef(lm(y~x))

beta1<-cor(y,x)*sd(y)/sd(x)
beta0<-mean(y) - beta1*mean(x)

rbind(c(beta0, beta1), coef(lm(y~x)))
```

## Fitting a linear model through a data

```{r Fitting a Linear model through the galton, fig.height=4,fig.width=4, echo=FALSE}
library(UsingR)
library(ggplot2)
data(galton)

y<-galton$child - mean(galton$child)
x<-galton$parent - mean(galton$parent)
# n<-galton$child
# m<-galton$parent

freq.data<- as.data.frame(table(x,y))
#freq.data_2<- as.data.frame(table(n,m))

names(freq.data)<-c('child', 'parent', 'freq')

freq.data$child<-as.numeric(as.character(freq.data$child))
freq.data$parent<-as.numeric(as.character(freq.data$parent))

new.freq.data<-filter(freq.data, freq.data$freq>0)

head(new.freq.data)

g<-ggplot(freq.data, aes(x=parent, y=child))
g<-g+ scale_size(range = c(2,20), guide = "none")
g<-g+ geom_point(color="grey40", aes(size=freq+20, show_guide=FALSE))
g<-g+ geom_point(aes(colour=freq, size=freq))
g<-g+ scale_colour_gradient(low = "lightblue", high = "white")
g<-g+ geom_smooth(method = "lm", formula = y~x)

g  
  #mse<-mean((y-beta*x)^2)
  #g<-g+ggtitle(paste("beta=", beta, "mse=", round(mse,3)))
```