---
title: "Linear Regression"
author: "Sandeep Anand"
date: "March 31, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Demonstation of the lm() function and Simple Linear Regression

### This part is contributed from data archive at Princeton Examples from http://data.princeton.edu/R/readingData.html 

```{r Reading the Data}
fpe<-read.table("http://data.princeton.edu/wws509/datasets/effort.dat", header = TRUE)
```

## Scatter Plot of Change based on Effort of the above Family Planning data


```{r chagne vs Effort, echo=FALSE}
plot(fpe$effort, fpe$change, pch=21, bg="green")
title("Scatterplot of change by Effort", col.main="#3366CC")
```

## Fitting a linear Model

> Note first that lm is a function, and we assign the result to an object that we choose to call lmfit (for linear model fit). This stores the results of the fit for later examination.

> The argument to lm is a model formula, which has the response on the left of the tilde ~ (read "is modeled as") and a Wilkinson-Rogers model specification formula on the right. R uses

  +	to combine elementary terms, as in A+B
  + :	for interactions, as in A:B
  *	for both main effects and interactions, so A*B = A+B+A:B


```{r fitting a linear model, echo=TRUE}
lmfit<-lm( change ~ setting+effort, data = fpe)
summary(lmfit)
anova(lmfit)
plot(lmfit)
```
