---
title: "Multi-Variate-regression"
author: "Sandeep Anand"
date: "April 15, 2017"
output: html_document
---

```{r setup, include=FALSE, highlight=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

---
## MutiVariate Regression

```{r MV-1,echo=TRUE}
require(datasets);data(swiss);require(ggplot2);require(GGally)
g = ggpairs(swiss, lower=list(continous="smooth"), wrap=c(geom_smooth(method = "loess")))
g

```

--- Calling 'lm'
+ Fitting all the variables

```{r Fit-1,echo=TRUE}
summary(lm(Fertility~., data=swiss))$coef
```

+ The -0.17 , means that there is an expected decrease of 0.17 in fertility for men , for every 1% increase in percentage of males involved in Agriculture , and this is the case with other variables remaining constant
+ Just to consider that the these are actually percentages 
+ R actually gives us the t-value for a t-test of H0: Beta(agri) = 0

--- Case when we only do a regression of the response Fertility and the variable Agriculture and not consider the other variables here
+ Fitting only Agriculture

```{r Fit-2,echo=TRUE}
summary(lm(Fertility~Agriculture, data=swiss))$coef
```

+ You can see for this case the trend of how fertility depends on Agriculture actually changes in a different direction, when keeping the other variables fixed
+ **Regression is a thinking process , basically which particular variables/Covariates to include and which to not**
+ If you see a NA , as the estimate at the end of the coefficients that is probably a completely unnecessary and is not having any causal relationship
+ If a particular predictor/covariate is highly causal , then excluding that will cause a high residue of that to be used as the other covariate and the relatioship/coefficients of the linear model will change a lot, The response could pick up a big residual effect of a covariate that is not picked and it picked up as the current response
+ We can also plot the residuals to find the relationships for them


```{r Fit-3,echo=TRUE}
data(InsectSprays);library(ggplot2)
g<-ggplot(data=InsectSprays, aes(y=count, x=spray, fill=spray))
g<-g+geom_violin(colour="black", size=2)
g<-g+xlab("Spray Type")+ylab("Insect Count")
g
```

+ The Intercept for the lm() function is basically the mean of the particular value that is omitted here, for the below case it is Spray A
+ We cannot force include spray A , otherwise if we try, R will drop one of them to NA
+ Though if we really want to drop the Intercept , we can use it and we will see all the sprays , but in this case , obviously there are no relative comparisions , you will see just the direct means as the estimates
+ **The reason we do a relative with intercept is so that we can compare different datasets and their relationships , for this case we would like to know how each spray is different from spray A and the probability of the t test that we are conducting , the t test being the (estimate diff)/(Standard error) . If we omit the intercept , then we are just comparing each of the Sprays with 0, that is if they killed more number of intsects
```{r Fit-4, echo=TRUE}
summary(lm(count~spray, data=InsectSprays))$coef

# Do the same thing manually or hardcoded
summary(lm(count~
             I(1*(spray=="B"))+I(1*(spray=="C"))+I(1*(spray=="D")) + I(1*(spray=="E"))+  I(1*(spray=="F")) 
           , data = InsectSprays))$coef
summary(lm(count~spray-1,data=InsectSprays))$coef

```
---
Reordering the levels
+ Intercept is treated as the mean of that reference level
+ All the other sprays are comparision of that spray with the reference level
+ The tests will be test of the comparision with the spray A
```{r Fit-5, echo=TRUE}
spray2<-relevel(InsectSprays$spray, "C")
summary(lm(count~spray2,data=InsectSprays))$coef
```
--- Ancova
+ Analysis of Covariance
+ Fitting a number of lines with different intercepts 
+ If you multiply you will get an interaction
```{r Analysis of Covariance, echo=TRUE}
library(datasets);data(swiss);library(ggplot2)
head(swiss)

# Create a binary  catholic variable, 1 if the county is majority catholic and majority protestant
library(dplyr)
swiss = mutate(swiss, CatholicBin=1*(Catholic>50))

#fit4<- lm(Fertility~Agriculture, data=swiss)

#fit5<-lm(Fertility~Agriculture+factor(CatholicBin), data=swiss)

fit6<-lm(Fertility~Agriculture*factor(CatholicBin), data=swiss)

g<-ggplot(swiss, aes(x=Agriculture, y=Fertility, colour=factor(CatholicBin)))
g=g+geom_point(size=6, colour="black")+geom_point(size=4)
g=g+xlab("% in Agriculture") + ylab("Fertility")

g3<-g
g3<-g3+geom_abline(intercept = coef(fit6)[1], slope=coef(fit6)[2], size=2)
g3<-g3+geom_abline(intercept = coef(fit6)[1]+coef(fit6)[3], slope=coef(fit6)[2]+coef(fit6)[4], size=2)
g3
```