---
title: "Statistical Inference Project Part01"
author: "Sandeep Anand"
date: "March 26, 2017"
output: pdf_document
---

```{r setup, include=FALSE, highlight=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Project Heading

The project consists of two parts:

1. A simulation exercise.
2. Basic inferential data analysis.

### Starting with the Simulation Excercise

**Exponential Function**

* In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. 
* The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. 
* The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda.

**To-Do's [Step-by-Step]**

+ Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials. 
+ Show the sample mean and compare it to the theoretical mean of the distribution.
+ Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
+ Show that the distribution is approximately normal.

```{r Exponential function}
lambda<-0.2
n=40
B<-1000

samplev<- rep(NA,1000)

## Sample Vector
for (i in 1:1000){
  gen <- rexp(n,lambda)
  samplev[i] <- mean(gen)
}

```

## Showing the Sample Mean and Variance and Comparisions between the Sample and the Theoretical distributions


```{r comparision of Mean and Variance, echo=TRUE}
library(ggplot2)

#########################
#       Mean            #
#########################

sample_mean<-mean(samplev)

given_mean<-1/lambda

vect <- c(sample_mean, given_mean)

sample_mean
given_mean

g<-ggplot(data.frame(vect), aes(seq_along(vect), vect)) +
  geom_bar(stat = "identity")+ 
  geom_hline(yintercept = sample_mean, color="red", lty=2) + 
  geom_hline(yintercept = given_mean, color="blue")

print(g)


#########################
#       Variance        #
#########################

sample_var<-var(samplev)

given_var<-(1/lambda)^2/n

vect_var <- c(sample_var, given_var)

sample_var
given_var

g2<-ggplot(data.frame(vect_var), aes(seq_along(vect_var), vect_var)) +
  geom_bar(stat = "identity")+
  geom_hline(yintercept = sample_var, color="red", lty=2) + 
  geom_hline(yintercept = given_var, color="blue")

print(g2)

```


### Check to show that the distribution is Normal

####A random variable is log-normally distributed if its log is a normally distributed random variable and it fits a bell curve


```{r Multiplot function, echo=FALSE}
# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```


```{r Check to show that the distribution is Normal, echo=TRUE}

# To See if the disctribution is approximately Normal

g1<-ggplot(data = data.frame(samplev), aes(samplev))+
    geom_histogram(color="lightblue",bins = 20)+ggtitle("Check for a Normal Distribution")

g3<-ggplot(data=data.frame(samplev),aes(samplev))+
    geom_density(color="blue")+ggtitle("Check to see if this is a Bell Curve")

multiplot(g1, g3)

```

# Conclusion
**Result : The Distribution is Normal as seen from the Curves above**