---
title: "Statistical Inference Project Part02"
author: "Sandeep Anand"
date: "March 27, 2017"
output: pdf_document
---

## Part-2 Of the Statistical Inference Project: 
* Analyze the ToothGrowth data in the R datasets package
* Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
* State the conclusions

```{r Basic structure & Summary of the ToothGrowth data, echo=TRUE}
data(ToothGrowth)
str(ToothGrowth)

# Summary
summary(ToothGrowth)
```

# Description of analysis with Graphs
### Toothgrowth comparision with respect to two groups OJ and VC 
```{r Analysis of dosage, echo=TRUE}
library(ggplot2)

g4<-ggplot(ToothGrowth, aes(x=factor(dose), y=len, fill=factor(dose))) +
    geom_boxplot()+facet_grid(.~supp)+
    ggtitle("Analyzing Toothgrowth data based on Type of Dose")

print(g4)

```

### Toothgrowth comparision with averages
```{r Analysis of dosage trends, echo=TRUE}

avg<-aggregate(ToothGrowth$len,by=list(dose=ToothGrowth$dose,supp=ToothGrowth$supp),FUN=mean)

avg

g5<-ggplot(ToothGrowth, aes(x=factor(dose), y=len)) +
    geom_point(aes(color=supp))+facet_grid(.~supp)+
    ggtitle("Analyzing Toothgrowth data trends based on Dosage")


print(g5)

g<- ggplot(avg, aes(x=factor(dose), y=avg$x, color=avg$supp))+
    geom_point(aes(size=avg$x))
print(g)

```


## Using Confidence Intervals and hypothesis tests to compare tooth growth by Supplement and Dosage.
```{r confidence Intervals and hypothesis tests, echo=TRUE}

print(t.test(len~supp,data = ToothGrowth, paired = FALSE))

```

# Conclusions

### The boxplots show that the growth is faster in case of dose = VC than dose = OJ . The 95% Confidence interval is got as (-0.171,7.571) 
As tooth size increases , dosage is also higher