---
title: "Starbucks Store Location finder and Density of Starbucks Store in the United States"
author: "Sandeep Anand"
date: "June 4, 2017"
output: html_document
params: 
  n:  100
  d:  !r Sys.Date()
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```


## Description 

###
- Load the Starbucks data from https://opendata.socrata.com/api/views/ddym-zvjk/rows.csv
- clean and Subset the data
- Use Leaflet to plot All the data on the Map for only California with Labels of the respective Stores, with Street Names
- Zoom In to find the Names and Labels of the stores present
- Making a heatmap using leaflet , of the stores present in the USA

```{r Load and clean the Data, echo=FALSE}
path <- "C:/Public/R-Projects/R-projects/starbucks"
file <- list.files(path = path)
finalpath <- file.path(path, file)
myData <- read.csv(file = finalpath, header = TRUE, sep = ",")
toselect<-c("Latitude", "Longitude" , "State","Name", "Street.Address", "City")
newData<-subset.data.frame(myData, select = toselect)
newData.CA<-subset.data.frame(newData, State=="CA", select = c("Latitude", "Longitude","Name", "Street.Address", "City"))
newData.CA$Information = paste("Starbucks",newData.CA$Name,",",newData.CA$Street.Address,",", newData.CA$City, ",", "CA")
```

## `r params$d`

```{r Fancy Map, echo=FALSE}
library(leaflet)
leaflet() %>% addTiles() %>% addMarkers(data=newData.CA, lat = ~Latitude, lng = ~Longitude, popup = newData.CA$Information, clusterOptions = markerClusterOptions())
```


## Density Plots with Circles

###
- Finding out where there are the maximum density of Starbucks stores present in the United States 
- Click on the Circles to know about the State

```{r Density Plot, echo=FALSE}
library(leaflet)
countvec <- rep(NA, 51)

State <- as.vector(unique(newData$State))

for (i in 1:length(State)){
  countvec[i] <- length(which(newData$State == State[i]))
}

# Column bind
newcombData <- cbind.data.frame(State, countvec)

# Getting the Data from the State information file
path2 <- "C:/Public/R-Projects/R-projects/statesinfo"
file2 <- list.files(path = path2)
finalpath2 <- file.path(path2, file2)
myDatastates <- read.csv(file = finalpath2, header = TRUE, sep = ",")

newmergedData <- merge(myDatastates, newcombData, by="State", all=TRUE)

m = leaflet(newmergedData) %>% addTiles() %>%
  addCircles(lng = ~Longitude, lat = ~Latitude, weight=1, radius = ~countvec*200, popup = ~City)
m

# # Creating a continous palette function
# pal = colorNumeric(
#   palette = "Greens",
#   domain = newmergedData$countvec
# )
# 
# coordinates(newmergedData) <- cbind(newmergedData$Longitude, newmergedData$Latitude)
# proj4string(newmergedData) <- CRS
# 
# n = leaflet(newmergedData) %>%
#   addPolygons(stroke = FALSE, 
#               smoothFactor = 0.2,
#               fillOpacity = 1,
#               color = ~pal(newmergedData$countvec),
#               popup = ~City)

```
